---
title: Review for Unit One # Title of the document

execute:               # Default settings for code blocks
  warning: false       # Do not print warnings
  message: false       # Do not print messages
  size: "scriptsize"   # Set code size to script size

# Below will do html by default. To render the entire project
# Use the following code:
# quarto::quarto_render("Review.qmd")
format:
  html:       # Specify HTML formatting options 
    number-sections: true      # Number Sections
    code-fold: true            # Button to show code
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables and figures
  pdf:        # Specify .pdf formatting options 
    documentclass: article     # Document Class
    number-sections: true      # Number Sections
    reference-section-title: "References" # Title for reference section
    code-line-numbers: true    # Line numbers printed in code
    cap-location: top          # Captions at the top of tables 
    geometry:
      - top=0.75in
      - left=0.75in
      - bottom=0.75in
      - right=0.75in
format-links: [pdf]
---
```{r}
#| echo: false

default_chunk_hook  <- knitr::knit_hooks$get("chunk")

latex_font_size <- c("Huge", "huge", "LARGE", "Large", 
                     "large", "normalsize", "small", 
                     "footnotesize", "scriptsize", "tiny")

knitr::knit_hooks$set(chunk = function(x, options) {
  x <- default_chunk_hook(x, options)
  if(options$size %in% latex_font_size) {
    paste0("\n \\", options$size, "\n\n", 
      x, 
      "\n\n \\normalsize"
    )
  } else {
    x
  }
})
knitr::opts_knit$set(root.dir = here::here())
```

In this document, we will review the content from Unit 1: Coding in R. 

# Calculations

R can be used as a calculator. To compute this expression: 
$4 * \left( 2 + \frac{2}{3} \right)$

you must type 

```{r}
(4 * (2 + (2/3)))
```

We can also perform integer division using %/%, and use the modulus operator %%.

For example, we will calculate $834 / 23$ using division, integer division, and the modulus. 

```{r}
(834/23)

(834 %/% 23)

(834 %% 23)
```

# Equalities and Logicals

Equalities and Logicals were essential parts of future units. 
Equalities commonly used were: Less than (<), Less than or equal to (<=), Greater than (>), Greater than or equal to (>=), Equal (==), and Not equal (!=).

Equalities are most commonly used to compare numbers, however they can also compare strings. 

The three logical operators are and (&), or (|), and not (!).
These operators have an order of operations: Not -> and -> or. Generally, Arithmetic is evaluated before comparisons, and comparisons are evaluated before logicals.

Here are some example uses of equalities with numbers:

```{r}
(4 > 3)
(3 > 3)
(3 >= 3)
(2 <= 3)
(3 == 3)
(3 != 5)
```

In the case of strings, a string is considered "greater" if its first letter comes later in the alphabet. If two strings are the same and one of them is capitalized, the capital string is "greater" than the lowercase one. Here are some examples of equalities being applied to strings:

```{r}
("hi" > "goodbye")
("hi" >= "goodbye")
("Hi" > "goodbye")
("hi" > "Hi")
("Hello" <= "bye")
```

# Packages

R has packages, which are downloadables that come with premade functions. Packages are extremely useful, however to use one, it must first be installed. This can be done as such:

```{r}
#| eval = FALSE
install.packages("tidyverse")
```

Once a package is installed, it must be loaded by calling library(). Install.packages is like downloading an app from the app store - only necessary once, while Library() is like pressing on the app to open it on your home screen - you have to do it every time to use it. The package will stay in use in your r session if opened once, so it is not necessary to call library() more than once with the same package in the same session.
```{r}
library(tidyverse)
```

To learn more about a specific package, help(), citation(), demo(), and vignette() can be called for a specific package by inserting the name of the package into the parenthesis like so.

```{r}
help("tidyverse")
```

# Objects in R

Values such as a string or a number can be assigned to an named object for easy reuse by using = or <-. For example:

```{r}
a = 5
(c=a*8)
(a >= c)
```

Each object in R has a class. The class of an object represents how R decides to apply a function towards it. To determine the class of an object, call class() as below.

```{r}
class(a)
```

There are five classes: Logical, integer, numeric, complex, and character. When comparing objects of different classes, whichever class appears furthest to the right in that list is "greater". For example: A string is greater than a number.

```{r}
(6>"6")
```

# Vectors 

A vector is a common data structure that can store multiple values. It can be created manually, using a sequence, or repeating values. Vectors are also atomic - they can only have one type of data. A vector takes the class of whichever object in it is furthest to the right in the list above. In this example, we create a vector manually. Note that the class is character, because "hi" is in the vector.

```{r}
vector1 = c(8, FALSE, "hi")
class(vector1)
```

Here, we create vectors using seq() and rep(). Note that rep() can repeat terms consecutively by using each, or it can repeat sequences using times. Note that repeatedvec is different based on if we use each or times.
```{r}
(evens = seq(from = 2, to = 20, by = 2))
(repeatedvec = rep(x = c(1, 45, "hi"), each = 2))
(repeatedvec = rep(x = c(1, 45, "hi"), times = 2))
```

A factor vector is a vector designed for categorical variables. There is a set of possible observations called "levels" that each part of a factor vector can take on. By default, these levels are set in alphabetical order, but they can be changed. Here is an example of a creation of a factor vector: 

```{r}
(teams = c("red sox", "celtics", "bruins", "patriots", "patriots", "patriots", "red sox", "bruins"))
factor(teams)
factor(teams, levels = c("red sox", "bruins", "celtics", "patriots"))
```

Note that the second time we used factor, since we specified the levels in order, the levels are now in that order.

Vectors can also be indexed to take a specific value out. Note that in R, indexing starts at 1 and not 0, which is different than some other programming languages. 

```{r}
teams = c("Celtics", "Red Sox", "Bruins", "Patriots")
teams[1]
```

To summarize a vector, the tools used depend on the type of vector. A numeric vector can be summarized using mean, median, standard deviation, interquartile range, skewness, or kurtosis. A factor vector can be summarized by the number of levels, most occuring level, or the proportion of each level.

# Matrices in R

A matrix is a rectangular object that has rows and columns. The rows and columns of a matrix are both vectors. Matrices are atomic, just like vectors. Matrices can also be indexed to get a specific item. 

```{r}
(matrix1 = matrix(data = c(1, 3, 4, 9), nrow = 3, ncol = 3))
```

Some basic linear algebra functions can be easily applied to matrices, such as creating a diagonal matrix, transposing, matrix multiplication, and calculating the determinant and eigenvalues. Importantly, using the regular * sign will componentwise multiply the matrix, and it is necessary to do %*% to perform regular matrix multiplication. 

```{r}
(matrix1 %*% matrix1)
(matrix1 * matrix1)
(det(matrix1))
(t(matrix1))
(eigen(matrix1)) 
```

# Lists

A list is similar to a vector, but it can contain objects of multiple types. To index a list, you use two brackets like so: list[[]]. 

# Data frames

A data frame is a list of vectors that are all the same size. Classes of columns need not be the same, but the elements in each column must be the same class. Data frames are created similar to matrices. 

```{r}
(df = data.frame(numbers = c(1, 5, 193), letters = c("A", "Y", "J")))
```

The contents of data frames can be viewed using view(), head(), and tail(). The dimension, row count, and column count can be accessed with dim(), nrow(), and ncol(). To index data frames, use single brackets. Multiple rows and columns can be pulled from a data frame. 

```{r}
(df[3, 2])
(df[c(1), ])
(df[, 1])
```

# Reading data frames

To read in data, we generally use the command read.csv, or read_csv (part of the tidyverse package). These commands only work for .csv files. While these two commands both work to read in csv files, read_csv is faster, and produces tibbles (a tidyverse alternative to data frames).

# Conditionals 

The most common use of conditionals is if/else. These come in two forms: shorthand, and block.

Shorthand:
```{r}
#| eval = FALSE
ifelse(condition, value when true, value when false)
```

Block:
```{r}
#| eval = FALSE
if(condition1){
    # do this
}else if(condition2){
    # do this instead
} else{
    # or do this 
}

```

Block notation is a more traditional way of creating conditionals (like in Python), and also does not have a limit on how many conditions can be passed in. However, block notation does not by default work on vectors, and therefore requires a loop for use on vectors.

# Loops

Loops essentially come in two different forms: for, and while. A for loop will complete a given task the number of times specified in the creation of the loop. A while loop will complete the tasks until the condition is met. While loops can be constructed using repeat or while. With repeat, you can choose whether the task should be completed before or after checking the condition. With while, the loop will only run while the condition is true and will immediately break out once the condition becomes false, which is checked at the beginning of each new iteration of the loop.  

Examples: 



```{r}
for (i in 1:4){
    print("This is a for loop.")
}


a = 0
repeat{
    print("hello")
    a = a + 1
    if(a > 3){
        break;
    }
}

a = 0
repeat{
    a = a + 1
    if(a > 3){
        break;
    }
    print("hi")
}

a = 0
while(a < 4){
    print("This is a while loop")
    a = a + 1
}

a = 0
while(a < 4){
    a = a + 1
    print("This is a while loop with different counter placement")
}
```

Note that in the repeat loops, the placement of the conditional can affect the amount of times the resulting message is printed. This is not the case in the while loops, where the output is printed four times regardless of placement of the counting variable.

# Functions

A function is a group of commands that must be defined once, and then can be used multiple times on different inputs. It is of the format

```{r}
#| eval = FALSE
function1 = function(arguements){
    #tasks
}
```

In the function definition, input defaults can be set. These are useful for complicated functions like ggplot, and they allow the user to use the function with minimal inputs. We can also choose an object to output at the end of the function, by using return or by calling the variable on the last line of the function. 

Here is an example function: 

```{r}
adding_entry = function(v){
    count = 0
    len = length(v)
    for (i in 1:len){
        print("I am increasing the length of the input by 1!")
        count = count + 1
        v[len+i] = count
    }
    v
}
adding_entry(c(1, 5, 9))
```

This function prints what v ends up being, because v is called at the end of the function. return(v) would have the same results. 

Functions can be applied to different things. apply() applies a function to rows or columns of a matrix, lapply() does the same to each element of a vector and returns a list, and sapply() does the same as lapply but returns a vector. 


